Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design

### 2025-04-07 src/main.tex
[text] answer florina's notes, and compression to fit on 4 pages

\newcommand{\fmc}[1]{{\color{magenta}#1}} % Florina Ciorba

@@

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' (for example, \df\ for Docker containers) made of a list of commands to execute from a ``base image'' (the \texttt{FROM} instruction in \df s).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
Administrators can also remove these base images, making all recipes un-buildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus \fmc{endangers} the reproducibility of the results.
One solution would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (for example, Zenodo~\cite{zenodo}).
But this comes with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

@@

In this paper, we propose our \textbf{initial design of a longitudinal study of the variations of the software environments produced by \df s from research artifacts.}
We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months. \fmc{Could we give here the period October 1st, 2024 - April 1st, 2025 of these 7 months?}
% \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}

@@

This section presents the different information collected by \ecg, and their motivations.

@@

In this section, we present the initial protocol of our longitudinal study.

>>>>

% \newcommand{\fmc}[1]{{\color{magenta}#1}} % Florina Ciorba
\newcommand{\fmc}[1]{{}} % Florina Ciorba

@@

One difficult aspect of creating a truly reproducible artifact is to properly control the software environment: packages and libraries versions.
Several works already shed light on the possible variability of results given a badly captured software environment~\cite{mytkowicz_producing_nodate, sokolowski2024impact}.
To address this issue, the conference \adae\ committees have advised authors to provide containers or virtual machines to share their artifacts in order to ``freeze'' their software environment~\cite{eurosys25, aetips, csartifacts}.
%However, authors generate their artifact right before the artifact submission~\cite{guilloteau2024longevity, guilloteau2024frustrations}.
In the case of usual virtualization techniques to generate containers or virtual machines, authors have to write the ``recipe'' (for example, \df\ for Docker containers) made of a list of commands to execute from a ``base image'' (the \texttt{FROM} instruction in \df s).
This approach has several limitations.
First, these base images can vary, either because of bad practices from the users (\eg\ using the \texttt{latest} version of the images) or from the administrators of the image (\eg\ pushing a different image to a previously defined tag).
Administrators can also remove these base images, making all recipes un-buildable~\cite{nvidia_cuda_lifetime}.
Another limitation of this approach is the heavy dependency on non-reproducible tools such as classical package managers (\eg\ \texttt{apt}, \texttt{dpkg}, \texttt{yum}).
Hence, rebuilding a recipe of a container or a virtual machine can yield different results from the authors' image, and thus endangers the reproducibility of the results.
One would be for the authors to store and then provide the built container itself (\ie tarball) on some long-term storage (\eg\ Zenodo~\cite{zenodo}).
But this comes with the downside of requiring more storage space than just storing the recipe (binary vs. text file), and thus has some \textbf{severe scalability and sustainability issues in terms of longevity}~\cite{monroe2023preservation, guilloteau2024longevity}.
Moreover, storing the built binary image exhibits limitations when needed to \emph{inspect} the actual content of the image, and when a modification of the software environment capture by this image is needed~\cite{mercier2018considering}.
Having the opportunity to extend previous scientific work is a crucial aspect in any scientific field.

@@

In this paper, we propose our \textbf{initial design of a longitudinal study of the variations of the software environments produced by \df s from research artifacts.}
We show initial results from five \df s extracted from the Euro-Par 2024 conference reproducibility artifacts built every month for seven months, from October 1$^{st}$ 2024 to April 1$^{st}$ 2025.
% \todo{The goal is not to dunk on authors or conferences, but simply to exhibit the potential issues with \df s.}

@@

This section presents the different information collected by \ecg.%, and their motivations.

@@

% In this section, we present the initial protocol of our longitudinal study.
This section presents the initial protocol of our longitudinal study.
